R Python, and Ruby clients for GBIF species occurrence data

### 2017-09-26 manuscript.tex
hopefully fixed issue #8

  Discussion. Programmatic acccess to GBIF will facilitate more open and
  reproducible science - the three GBIF clients described herein are a
  significant contribution towards this goal.
  \end{abstract}

@@

In Python, we can similary search for datasets. Here, search for
datasets of type \texttt{OCCURRENCE}:

@@

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (21534).

@@

The download service, like the occurrence search service, is
rate-limited. That is, you can only have one to three downloads running
simultaneously for your user credentials. However, simply check when a
download job is complete, then you can start a new download request. See
``Queueing Download Requests'' below for help automating many download
requests in R.

@@

The Ruby library \texttt{gbifrb} does not yet have occurrene download
functionality.

@@

In this example, we plot actual occurrence data for \emph{Bradypus}
species against a single predictor variable, BIO1 (annual mean
temperature). This is only ont step in a species distribution modelling
nworkflow.

@@

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{url <-}\StringTok{ 'https://raw.githubusercontent.com/amyxzhang/}
\StringTok{boundingbox-cities/master/boundbox.txt'}
\NormalTok{rawdat <-}\StringTok{ }\KeywordTok{content}\NormalTok{(}\KeywordTok{GET}\NormalTok{(}\KeywordTok{sub}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\StringTok{""}\NormalTok{, url)), }\DataTypeTok{as =} \StringTok{"text"}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.table}\NormalTok{(}
  \DataTypeTok{text =}\NormalTok{ rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }
  \DataTypeTok{sep =} \StringTok{"}\CharTok{\textbackslash{}t}\StringTok{"}\NormalTok{, }\DataTypeTok{col.names =} \KeywordTok{c}\NormalTok{(}\StringTok{"city"}\NormalTok{,}\StringTok{"minlat"}\NormalTok{,}\StringTok{"maxlon"}\NormalTok{,}\StringTok{"maxlat"}\NormalTok{,}\StringTok{"minlon"}\NormalTok{),}
  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
  \DataTypeTok{city =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{city, }\DataTypeTok{minlon =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{minlon, }
  \DataTypeTok{minlat =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{minlat, }\DataTypeTok{maxlon =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{maxlon, }
  \DataTypeTok{maxlat =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{maxlat,}
  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{getdata <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x)\{}
\NormalTok{  coords <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(x[}\KeywordTok{c}\NormalTok{(}\StringTok{'minlon'}\NormalTok{,}\StringTok{'minlat'}\NormalTok{,}\StringTok{'maxlon'}\NormalTok{,}\StringTok{'maxlat'}\NormalTok{)])}
\NormalTok{  wkt <-}\StringTok{ }\NormalTok{wicket}\OperatorTok{::}\KeywordTok{wkt_correct}\NormalTok{(wicket}\OperatorTok{::}\KeywordTok{bounding_wkt}\NormalTok{(}\DataTypeTok{values =}\NormalTok{ coords))}
\NormalTok{  num <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =}\NormalTok{ wkt, }\DataTypeTok{limit =} \DecValTok{1}\NormalTok{)}\OperatorTok{$}\NormalTok{meta}\OperatorTok{$}\NormalTok{count}
  \KeywordTok{data.frame}\NormalTok{(}
    \DataTypeTok{city =}\NormalTok{ x[}\StringTok{'city'}\NormalTok{], }
    \DataTypeTok{richness =}\NormalTok{ num, }
    \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}
\NormalTok{  )}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{'rgbif'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'raster'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'sp'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'sf'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'rgeos'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'scales'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Get GBIF Data for Quercus lobata}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{keyQl <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Quercus lobata'}\NormalTok{, }\DataTypeTok{kingdom=}\StringTok{'plants'}\NormalTok{)}\OperatorTok{$}\NormalTok{speciesKey}
\NormalTok{dat.Ql <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey=}\NormalTok{keyQl, }\DataTypeTok{return=}\StringTok{'data'}\NormalTok{, }\DataTypeTok{limit=}\DecValTok{50000}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Get Distribution map of Q. lobata Atlas of US Trees (Little, E.)}

From \url{http://esp.cr.usgs.gov/data/little}. And save shapefile in
same directory

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{url <-}\StringTok{ 'https://github.com/sckott/gbifms/blob/gh-pages/querloba.zip?raw=true'}
\NormalTok{tmp <-}\StringTok{ }\KeywordTok{tempdir}\NormalTok{()}
\KeywordTok{download.file}\NormalTok{(url, }\DataTypeTok{destfile =} \StringTok{"querloba.zip"}\NormalTok{)}
\KeywordTok{unzip}\NormalTok{(}\StringTok{"querloba.zip"}\NormalTok{, }\DataTypeTok{exdir =}\NormalTok{ tmp)}
\NormalTok{ql <-}\StringTok{ }\NormalTok{sf}\OperatorTok{::}\KeywordTok{read_sf}\NormalTok{(}\KeywordTok{file.path}\NormalTok{(tmp, }\StringTok{"querloba/querloba.shp"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{alt.USA <-}\StringTok{ }\KeywordTok{getData}\NormalTok{(}\StringTok{'alt'}\NormalTok{, }\DataTypeTok{country =} \StringTok{'USA'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Create Hillshade of US}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{alt.USA <-}\StringTok{ }\NormalTok{alt.USA[[}\DecValTok{1}\NormalTok{]]}
\NormalTok{slope.USA <-}\StringTok{ }\KeywordTok{terrain}\NormalTok{(alt.USA, }\DataTypeTok{opt =} \StringTok{'slope'}\NormalTok{)}
\NormalTok{aspect.USA <-}\StringTok{ }\KeywordTok{terrain}\NormalTok{(alt.USA, }\DataTypeTok{opt =} \StringTok{'aspect'}\NormalTok{)}
\NormalTok{hill.USA <-}\StringTok{ }\KeywordTok{hillShade}\NormalTok{(slope.USA, aspect.USA, }\DataTypeTok{angle =} \DecValTok{45}\NormalTok{, }\DataTypeTok{direction =} \DecValTok{315}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(hill.USA, }\DataTypeTok{col =} \KeywordTok{grey}\NormalTok{(}\DecValTok{0}\OperatorTok{:}\DecValTok{100}\OperatorTok{/}\DecValTok{100}\NormalTok{), }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{,}
     \DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(}\OperatorTok{-}\DecValTok{125}\NormalTok{, }\OperatorTok{-}\DecValTok{116}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{32}\NormalTok{, }\DecValTok{42}\NormalTok{),}
     \DataTypeTok{main =} \StringTok{'Distribution of Quercus lobata'}\NormalTok{,}
     \DataTypeTok{xlab =} \StringTok{"Long"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{'Lat'}\NormalTok{)}
\CommentTok{# add shape from Atlas of US Trees}
\KeywordTok{plot}\NormalTok{(ql, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.6}\NormalTok{),}
     \DataTypeTok{border =} \OtherTok{FALSE}\NormalTok{)}
\CommentTok{# add Gbif presence points}
\KeywordTok{points}\NormalTok{(dat.Ql}\OperatorTok{$}\NormalTok{decimalLongitude, dat.Ql}\OperatorTok{$}\NormalTok{decimalLatitude,}
       \DataTypeTok{cex =}\NormalTok{ .}\DecValTok{7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \OperatorTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{40.5}\NormalTok{, }\StringTok{"GBIF Data"}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{,}
       \DataTypeTok{col =} \StringTok{'darkgreen'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{, }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =}\NormalTok{ .}\DecValTok{8}\NormalTok{)}
\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \OperatorTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{41.5}\NormalTok{,}
       \DataTypeTok{legend =} \StringTok{"Atlas of United States Trees }\CharTok{\textbackslash{}n}\StringTok{ (Little, E. 1971)"}\NormalTok{,}
       \DataTypeTok{pt.cex =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex =}\NormalTok{ .}\DecValTok{8}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'white'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{components/figure/manuscript-unnamed-chunk-65-1.pdf}

>>>>

  Discussion. Programmatic access to GBIF will facilitate more open and
  reproducible science - the three GBIF clients described herein are a
  significant contribution towards this goal.
  \end{abstract}

@@

In Python, we can similarly search for datasets. Here, search for
datasets of type \texttt{OCCURRENCE}:

@@

Dataset metrics are another useful way of figuring out what datasets you
may want to use. One drawback is that these metrics data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (21687).

@@

The download service, like the occurrence search service, is
rate-limited. That is, you can only have one to three downloads running
simultaneously for your user credentials. However, simply check when a
download job is complete, then you can start a new download request. See
``Queuing Download Requests'' below for help automating many download
requests in R.

@@

The Ruby library \texttt{gbifrb} does not yet have occurrence download
functionality.

@@

In this example, we plot actual occurrence data for \emph{Bradypus}
species against a single predictor variable, BIO1 (annual mean
temperature). This is only one step in a species distribution modelling
workflow.

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{devtools}\OperatorTok{::}\KeywordTok{install_github}\NormalTok{(}\StringTok{"ropensci/wicket"}\NormalTok{, }\DataTypeTok{force =} \OtherTok{TRUE}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{url <-}\StringTok{ 'https://raw.githubusercontent.com/amyxzhang/}
\StringTok{boundingbox-cities/master/boundbox.txt'}
\NormalTok{rawdat <-}\StringTok{ }\KeywordTok{content}\NormalTok{(}\KeywordTok{GET}\NormalTok{(}\KeywordTok{sub}\NormalTok{(}\StringTok{"}\CharTok{\textbackslash{}n}\StringTok{"}\NormalTok{, }\StringTok{""}\NormalTok{, url)), }\DataTypeTok{as =} \StringTok{"text"}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.table}\NormalTok{(}
  \DataTypeTok{text =}\NormalTok{ rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{,}
  \DataTypeTok{sep =} \StringTok{"}\CharTok{\textbackslash{}t}\StringTok{"}\NormalTok{, }\DataTypeTok{col.names =} \KeywordTok{c}\NormalTok{(}\StringTok{"city"}\NormalTok{,}\StringTok{"minlat"}\NormalTok{,}\StringTok{"maxlon"}\NormalTok{,}\StringTok{"maxlat"}\NormalTok{,}\StringTok{"minlon"}\NormalTok{),}
  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}
  \DataTypeTok{city =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{city, }\DataTypeTok{minlon =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{minlon,}
  \DataTypeTok{minlat =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{minlat, }\DataTypeTok{maxlon =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{maxlon,}
  \DataTypeTok{maxlat =}\NormalTok{ dat}\OperatorTok{$}\NormalTok{maxlat,}
  \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}
\NormalTok{)}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{getdata <-}\StringTok{ }\ControlFlowTok{function}\NormalTok{(x)\{}
\NormalTok{  coords <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(x[}\KeywordTok{c}\NormalTok{(}\StringTok{'minlon'}\NormalTok{,}\StringTok{'minlat'}\NormalTok{,}\StringTok{'maxlon'}\NormalTok{,}\StringTok{'maxlat'}\NormalTok{)])}
\NormalTok{  wkt <-}\StringTok{ }\NormalTok{wicket}\OperatorTok{::}\KeywordTok{wkt_correct}\NormalTok{(wicket}\OperatorTok{::}\KeywordTok{bounding_wkt}\NormalTok{(}\DataTypeTok{values =}\NormalTok{ coords))}
\NormalTok{  num <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =}\NormalTok{ wkt, }\DataTypeTok{limit =} \DecValTok{1}\NormalTok{)}\OperatorTok{$}\NormalTok{meta}\OperatorTok{$}\NormalTok{count}
  \KeywordTok{data.frame}\NormalTok{(}
    \DataTypeTok{city =}\NormalTok{ x[}\StringTok{'city'}\NormalTok{],}
    \DataTypeTok{richness =}\NormalTok{ num,}
    \DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}
\NormalTok{  )}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{'rgbif'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'raster'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'sp'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'sf'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'rgeos'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'scales'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'rnaturalearth'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Get GBIF Data for Fraxinus excelsior}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{keyFe <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name =} \StringTok{'Fraxinus excelsior'}\NormalTok{, }\DataTypeTok{kingdom =} \StringTok{'plants'}\NormalTok{)}\OperatorTok{$}\NormalTok{speciesKey}
\NormalTok{dat.Fe <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =}\NormalTok{ keyFe, }\DataTypeTok{return =} \StringTok{'data'}\NormalTok{, }\DataTypeTok{limit =}\NormalTok{ 10000L)}
\end{Highlighting}
\end{Shaded}

\emph{Get Distribution map of F. excelsior European Forest Genetic
Resources Programme}

From \url{http://www.euforgen.org/species/fraxinus-excelsior/}. And save
shapefile in same directory

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{url <-}\StringTok{ 'http://www.euforgen.org/fileadmin/templates/euforgen.org/upload/Documents/Maps/Shapefile/Fraxinus_excelsior.zip'}
\NormalTok{tmp <-}\StringTok{ }\KeywordTok{tempdir}\NormalTok{()}
\KeywordTok{download.file}\NormalTok{(url, }\DataTypeTok{destfile =} \StringTok{"fraxinus_excelsior.zip"}\NormalTok{)}
\KeywordTok{unzip}\NormalTok{(}\StringTok{"fraxinus_excelsior.zip"}\NormalTok{, }\DataTypeTok{exdir =}\NormalTok{ tmp)}
\NormalTok{fe <-}\StringTok{ }\NormalTok{sf}\OperatorTok{::}\KeywordTok{read_sf}\NormalTok{(}\KeywordTok{file.path}\NormalTok{(tmp, }\StringTok{"Fraxinus_excelsior_EUFORGEN.shp"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{eur <-}\StringTok{ }\NormalTok{rnaturalearth}\OperatorTok{::}\KeywordTok{ne_countries}\NormalTok{(}\DataTypeTok{continent =} \StringTok{"europe"}\NormalTok{, }\DataTypeTok{type =} \StringTok{"map_units"}\NormalTok{)}
\NormalTok{eur1 <-}\StringTok{ }\NormalTok{eur[eur}\OperatorTok{$}\NormalTok{sovereignt }\OperatorTok{!=}\StringTok{ "Russia"}\NormalTok{, ]}
\end{Highlighting}
\end{Shaded}

@@

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(eur1, }\DataTypeTok{col =} \StringTok{"darkgrey"}\NormalTok{, }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{,}
     \DataTypeTok{main =} \StringTok{'Distribution of Fraxinus excelsior'}\NormalTok{)}
\CommentTok{# add distribution range layer}
\KeywordTok{plot}\NormalTok{(fe, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.5}\NormalTok{), }\DataTypeTok{border =} \OtherTok{FALSE}\NormalTok{)}
\CommentTok{# add Gbif presence points}
\KeywordTok{points}\NormalTok{(dat.Fe}\OperatorTok{$}\NormalTok{decimalLongitude, dat.Fe}\OperatorTok{$}\NormalTok{decimalLatitude,}
       \DataTypeTok{cex =}\NormalTok{ .}\DecValTok{7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \DecValTok{38}\NormalTok{, }\DataTypeTok{y =} \DecValTok{81}\NormalTok{, }\KeywordTok{c}\NormalTok{(}\StringTok{"GBIF Data"}\NormalTok{, }\StringTok{"Range Layer"}\NormalTok{), }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{bg =} \StringTok{"grey"}\NormalTok{,}
       \DataTypeTok{col =} \KeywordTok{c}\NormalTok{(}\StringTok{'darkgreen'}\NormalTok{, }\KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.5}\NormalTok{)), }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =}\NormalTok{ .}\DecValTok{8}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\includegraphics{components/figure/manuscript-unnamed-chunk-64-1.pdf}

### 2017-09-26 use-cases/oak-occurrences.Rmd
hopefully fixed issue #8

```{r}
library('rgbif')
library('raster')
library('sp')
library('maptools')
library('rgeos')
library('scales')
```

_Get GBIF Data for Quercus lobata_

```{r cache=TRUE}
keyQl <- name_backbone(name='Quercus lobata', kingdom='plants')$speciesKey
dat.Ql <- occ_search(taxonKey=keyQl, return='data', limit=50000)
```

_Get Distribution map of Q. lobata Atlas of US Trees (Little, E.)_

From [http://esp.cr.usgs.gov/data/little/](http://esp.cr.usgs.gov/data/little/). And save shapefile in same directory

```{r}
url <- 'http://esp.cr.usgs.gov/data/little/querloba.zip'
tmp <- tempdir()
download.file(url, destfile = "~/querloba.zip")
unzip("~/querloba.zip", exdir = tmp)
ql <- readShapePoly(file.path(tmp, "querloba.shp"))
```

@@

```{r}
alt.USA <- getData('alt', country = 'USA')
```

_Create Hillshade of US_

```{r}
alt.USA <- alt.USA[[1]]
slope.USA <- terrain(alt.USA, opt = 'slope')
aspect.USA <- terrain(alt.USA, opt = 'aspect')
hill.USA <- hillShade(slope.USA, aspect.USA, angle = 45, direction = 315)
```

@@

```{r fig.keep='last'}
plot(hill.USA, col = grey(0:100/100), legend = FALSE, xlim = c(-125, -116), ylim = c(32, 42), main = 'Distribution of Quercus lobata', xlab = "Long", ylab = 'Lat')
# add shape from Atlas of US Trees
plot(ql, add = TRUE, col = alpha("white", 0.6), border = FALSE)
# add Gbif presence points
points(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude, cex = .7, pch = 19, col = alpha("darkgreen", 0.8))
legend(x = -121, y = 40.5, "GBIF Data", pch = 19, col = 'darkgreen', bty = 'n', pt.cex = 1, cex = .8)
legend(x = -121, y = 41.5, "Atlas of United States Trees \n (Little, E. 1971)", pt.cex = 1.5, cex = .8, pch = 19, col = 'white', bty = 'n')
```

>>>>

```{r}
library('rgbif')
library('raster')
library('sp')
library('sf')
library('rgeos')
library('scales')
library('rnaturalearth')
```

_Get GBIF Data for Fraxinus excelsior_

```{r cache=TRUE}
keyFe <- name_backbone(name = 'Fraxinus excelsior', kingdom = 'plants')$speciesKey
dat.Fe <- occ_search(taxonKey = keyFe, return = 'data', limit = 10000L)
```

_Get Distribution map of F. excelsior European Forest Genetic Resources Programme_

From <http://www.euforgen.org/species/fraxinus-excelsior/>. And save shapefile in same directory

```{r cache = TRUE}
url <- 'http://www.euforgen.org/fileadmin/templates/euforgen.org/upload/Documents/Maps/Shapefile/Fraxinus_excelsior.zip'
tmp <- tempdir()
download.file(url, destfile = "fraxinus_excelsior.zip")
unzip("fraxinus_excelsior.zip", exdir = tmp)
fe <- sf::read_sf(file.path(tmp, "Fraxinus_excelsior_EUFORGEN.shp"))
```

@@

```{r cache = TRUE}
eur <- rnaturalearth::ne_countries(continent = "europe", type = "map_units")
eur1 <- eur[eur$sovereignt != "Russia", ]
```

@@

```{r fig.keep='last'}
plot(eur1, col = "darkgrey", legend = FALSE,
     main = 'Distribution of Fraxinus excelsior')
# add distribution range layer
plot(fe, add = TRUE, col = alpha("white", 0.5), border = FALSE)
# add Gbif presence points
points(dat.Fe$decimalLongitude, dat.Fe$decimalLatitude,
       cex = .7, pch = 19, col = alpha("darkgreen", 0.8))
legend(x = 38, y = 81, c("GBIF Data", "Range Layer"), pch = 19, bg = "grey",
       col = c('darkgreen', alpha("white", 0.5)), pt.cex = 1, cex = .8)
```
